Supplementary Information to "Improving Models to Predict Holocellulose and Klason Lignin Contents for Peat Soil Organic Matter with Mid Infrared Spectra

### 2022-07-25 analysis/paper/008-paper-supplementary.Rmd
Update `renv.lock`, update figure format in supporting info for manuscript 1. Update the text. Update the readme.
Address reviewer comments to manuscript 1.

```{r sup-setup, echo=FALSE}
knitr::opts_chunk$set(fig.align = "center", 
                      fig.pos = "H",
                      echo = FALSE,
                      cache = TRUE,
                      dev = "png")
```

@@

Throughout the main text, we analyzed weaknesses of the models as developed by @Hodgkins.2018 using a Bayesian implementation of these original models. Here, we show that working within a Bayesian framework does not create the bias or any other weaknesses we discussed. To this end, we compare predictions of the original non-Bayesian models (implemented with `lm` [@RCoreTeam.2020]) with those of the original Bayesian model we used as baseline model throughout the main text.  
Figure \@ref(fig:res-sup-original-models) compares median fitted values, and lower and upper 90% prediction intervals for both models.

@@

Here, we present results of a simulation study that shows exemplarily how changes in heights of other peaks can affect the normalized peak height of the `carb` peak and absorbances at other wavenumbers. With this, we want to demonstrate that both the original and improved models are sensitive to the overall spectra and therefore differences between training data and sample data can cause differences in predictions.  
In the simulation study, we simulated spectra with the same `carb` peak height, but a differently large OH-peak height. We used the script from @Hodgkins.2018 to compute the normalized `carb` peak height in the spectra and analyzed also the absorbance at other wavenumbers. Figure \@ref(fig:res-sup-spectra-sim-p) shows the simulated spectra.

@@

```{r res-sup-spectra-sim-p, echo=FALSE, out.width="80%", fig.height=7, fig.width=7, fig.cap="Results of the simulation experiment. (a) Plots of the simulated spectra with different OH-peak heights (columns) and different carb peak heights (colour gradients). Due to the normalization (division of each absorbance value by the sum of all absorbance value), the normalized simulated carb peak heights get smaller as the OH-peak gets larger. For this reason, the extracted normalized carb peak height depends not only on the simulated carb peak height, but is also sensitive to the OH-peak height (b). This is also true for absorbance values extracted at specific wavenumbers (c)."}
spectra_sim_p
```

@@

Figure \@ref(fig:res-p-gaussian-beta-depth-profile) shows depth profiles of holocellulose and Klason lignin contents for the peat cores from @Hodgkins.2018, predicted either with the original normal regression model or the beta regression model with the same model structure.

@@

Figure \@ref(fig:res-sup-kl-hol) summarizes the _measured_ holocellulose and Klason lignin contents in the OM samples of the training data set of @Hodgkins.2018.

```{r sup-kl-hol, echo=FALSE}
p1 <-
  d %>%
  ggplot(aes(y = (hol + kl) * 100, x = sample_type)) +
  geom_boxplot() +
  coord_flip() +
  labs(y = "Sample type", x = "Holocellulose + Klason lignin [mass-%]")

@@

```{r res-sup-kl-hol, echo=FALSE, out.width="100%", fig.height=4.5, fig.width=7, fig.cap='Measured holocellulose and Klason lignin contents in OM samples of the training data of \\citet{Hodgkins.2018}. A: Boxplots of the sum of the measured holocellulose and Klason lignin contents for different sample types in the training data. B: Klason lignin content versus holocellulose content for the training data samples. Colours represent different sample types and lines are linear regression lines fitted to the data.'}
p1 + p2 +
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
```

@@

Figure \@ref(fig:res-p-gaussian-beta-depth-profile-best) shows holocellulose and Klason lignin contents for the peat core samples from @Hodgkins.2018 predicted by the original model, the improved model using binned spectra, and the improved model using extracted peaks.

@@

Figure \@ref(fig:res-p-y-yhat-best) shows measured holocellulose and Klason lignin contents versus the respective fitted values of the original model, the improved model using binned spectra, and the improved model using extracted peaks.

```{r res-p-y-yhat-best, echo=FALSE, out.width="70%", fig.height=6.5, fig.width=6, fig.cap='Measured values of the training dataset from \\cite{Hodgkins.2018} versus fitted values of the original model (first column) in comparison to the improved models (second and third columns) for (a) holocellulose and (b) Klason lignin. Points are median predictions from the models, smooth lines and shaded regions are LOESS smoothers and corresponding 95\\% confidence intervals. The diagonal lines indicates ideal fit. Points above the line indicate overestimation by the model and points under the diagonal underestimation.'}
p_y_yhat_best
```

@@

Figure \@ref(fig:res-p-pred-diff-hol-spectra-diff2) shows MIRS for the training data, vegetation samples, and peat samples from @Hodgkins.2018, classified according to different `carb` peak heights and colored according to the prediction differences between the original model and the improved model using binned spectra for holocellulose. This plot makes visible the differences between the spectra for peat (and vegetation samples) in comparison to the training data.

@@

Figure \@ref(fig:res-p-pred-diff-hol-intensity2) shows prediction differences between the original model and the improved model using binned spectra for holocellulose in dependency of absorbance values at different wavenumbers and for different classes of `carb` peak heights.

@@

Figure \@ref(fig:res-p-pred-diff-kl-residuals) shows residuals of the original model, the improved model using binned spectra, and the improved model using extracted peaks versus standardized `arom15arom16` values and for different levels of the relative contribution of the `arom15` peak height to the `arom15arom16` value.

@@

Figure \@ref(fig:res-p-pred-diff-kl-histogram1) shows `arom15arom16` values for the training data, peat, and vegetation samples from @Hodgkins.2018, for different levels of the relative contribution of the `arom15` peak height to the `arom15arom16` value. Conditional on this contribution, peat samples regularly have larger or smaller `arom15arom16` values.

@@

Figure \@ref(fig:res-p-pred-diff-train2) shows measured holocellulose and Klason lignin contents versus the respective fitted values of the original model, the improved model using binned spectra, and the improved model using extracted peaks, while differentiating the sample type classes defined in @Hodgkins.2018. This makes visible that for Klason lignin, the two classes of training samples in the main text comprise different sample types (class 1: wood samples, type 2: leaves and grasses).

@@

Figure \@ref(fig:res-p-y-yhat-minerals) shows measured holocellulose and Klason lignin contents versus the respective fitted values of the original model, the improved model using binned spectra, the improved model using extracted peaks, the improved model using binned spectra trained with old magazine samples, and the improved model using extracted peaks trained with old magazine samples.

```{r res-p-y-yhat-minerals, echo=FALSE, out.width="100%", fig.height=3.5, fig.width=7.5, fig.cap='Predicted median holocellulose contents versus measured holocellulose contents for the Gaussian model with original model structure (first panel), the improved best models for both approaches (model including all peaks from the peak extraction procedure of \\cite{Hodgkins.2018}, model using binnes spectra with a bin width of 50 cm$^{-1}$), but fitted only to samples without mineral interferences (second and third panel), and the same models fitted also to samples with mineral interferences (last two panels). The diagonal line represent identical fitted and measured values. Colours differentiate samples with (blue) and without (red) mineral interferences.'}
p_y_yhat_minerals
```

@@

Figure \@ref(fig:res-p-oh-height-difference) shows differences in the absorbance at 3400 cm$^{-1}$ in area normalized spectra between the training, peat, and vegetation data.

>>>>

```{r sup-setup, echo=FALSE}
knitr::opts_chunk$set(fig.align = "center", 
                      fig.pos = "H",
                      echo = FALSE,
                      cache = TRUE,
                      dev = "pdf")
```

@@

Throughout the main text, we analyzed weaknesses of the models as developed by @Hodgkins.2018 using a Bayesian implementation of these original models. Here, we show that working within a Bayesian framework does not create the bias or any other weaknesses we discussed. To this end, we compare predictions of the original non-Bayesian models (implemented with `lm` [@RCoreTeam.2020]) with those of the original Bayesian model we used as a baseline model throughout the main text.  
Figure \@ref(fig:res-sup-original-models) compares median fitted values, and lower and upper 90% prediction intervals for both models.

@@

Here, we present results of a simulation study that shows exemplarily how changes in heights of other peaks can affect the normalized peak height of the `carb` peak and absorbances at other wavenumbers. With this, we want to demonstrate that both the original and improved models are sensitive to the overall spectra and therefore differences between training data and sample data can cause differences in predictions.  
In the simulation study, we simulated spectra with the same `carb` peak height, but a differently large OH-peak height. We used the script from @Hodgkins.2018 to compute the normalized `carb` peak height in the spectra and analyzed also the absorbance at other wavenumbers. Figure \@ref(fig:res-sup-spectra-sim-p) shows the simulated spectra.  
The implication of the simulation is that, when the spectra are normalized by dividion by the sum of all absorbance values in each spectrum, all spectral variables which we use for interpretation or prediction (e.g. by selecting the `carb` peak, or a bin at a specific wavenumber) are dependent on all other spectral variables in the spectra, even if we do not include them in the model.

@@

```{r res-sup-spectra-sim-p, echo=FALSE, out.width="80%", fig.height=7, fig.width=7, fig.cap="Results of the simulation experiment. (a) Plots of the simulated spectra with different OH-peak heights (columns) and different \\texttt{carb} peak heights. In this panel, the colour gradient represents different \\texttt{carb} peak heights. (b) Due to the normalization (division of each absorbance value by the sum of all absorbance values), the normalized simulated \\texttt{carb} peak heights (y axis values) get smaller as the OH-peak gets larger (colour gradient). For this reason, the extracted normalized \\texttt{carb} peak height depends not only on the simulated \\texttt{carb} peak height, but is also sensitive to the OH-peak height. (c) This can also be seen when, for these simulated spectra, absorbance values are extracted at specific wavenumbers, e.g. 1250 cm$^{-1}$."}
spectra_sim_p
```

@@

@@

```{r sup-kl-hol, echo=FALSE}
p1 <-
  d %>%
  ggplot(aes(y = (hol + kl) * 100, x = sample_type)) +
  geom_boxplot() +
  coord_flip() +
  labs(x = "Sample type", y = "Holocellulose + Klason lignin [mass-%]")

@@

```{r res-sup-kl-hol, echo=FALSE, out.width="100%", fig.height=4.5, fig.width=7, fig.cap='Measured holocellulose and Klason lignin contents in OM samples of the training data of \\citet{Hodgkins.2018}. A: Boxplots of the sum of the measured holocellulose and Klason lignin contents for different sample types in the training data. B: Klason lignin content versus holocellulose content for the training data samples. Colours represent different sample types and lines are linear regression lines fitted to the data.'}
p1 + p2 +
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") +
  plot_layout(guides = "collect", widths = c(1, 0.6)) &
  theme(legend.position = "bottom")
```

@@

@@

```{r res-p-y-yhat-best, echo=FALSE, out.width="70%", fig.height=6.5, fig.width=6, fig.cap='Measured values of the training dataset from \\cite{Hodgkins.2018} versus fitted values of the original model (first column) in comparison to the improved models (second and third columns) for (a) holocellulose and (b) Klason lignin. Points are median predictions from the models, smooth lines and shaded regions are LOESS smoothers and corresponding 95\\% confidence intervals. The diagonal lines indicates ideal fit. Points above the line indicate underestimation by the model and points under the diagonal overestimation.'}
p_y_yhat_best
```

@@

@@

@@

@@

@@

@@

```{r res-p-y-yhat-minerals, echo=FALSE, out.width="100%", fig.height=3.5, fig.width=7.5, fig.cap='Measured holocellulose contents versus predicted median holocellulose contents for the Gaussian model with original model structure (first panel), the improved best models for both approaches (model including all peaks from the peak extraction procedure of \\cite{Hodgkins.2018}, model using binned spectra with a bin width of 20 cm$^{-1}$), but fitted only to samples without mineral interferences (second and third panel), and the same models fitted also to samples with mineral interferences (last two panels). The diagonal lines represent identical measured and fitted values. Colours differentiate samples with (blue) and without (red) mineral interferences.'}
p_y_yhat_minerals
```

@@
